---
title: "Iteration_and_listcols"
author: "Bohan Zhu"
date: "2025-10-28"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(rvest)
```

## Make a list

```{r}
l =
  list(vec_numeric = 1:23,
       char_vec = c("Jeff"),
       mat = matrix(1:8, nrow = 2, ncol = 4),
       summary = summary(rnorm(1000, mean = 4))
       )

l

l[["vec_numeric"]]

l$mat
```

## Make a different list

```{r}
list_normals = 
  list(
    a = rnorm(30, mean = 3, sd = 1),
    b = rnorm(30, mean = 30, sd = 1),
    c = rnorm(30, mean = 3, sd = 10),
    d = rnorm(30, mean = -3, sd = 4)
  )
```


```{r}
mean_and_sd = function(x) {
  
  if (!is.numeric(x)) {
    stop("Argument x should be numeric")
  } else if (length(x) == 1) {
    stop("Cannot be computed for length 1 vectors")
  }
  
  mean_x = mean(x)
  sd_x = sd(x)

  tibble(
    mean = mean_x, 
    sd = sd_x
  )
}
```

```{r}
mean_and_sd(list_normals[[1]])
mean_and_sd(list_normals[[2]])
mean_and_sd(list_normals[[3]])
mean_and_sd(list_normals[[4]])
```

Use a loop to iterate!!

```{r}
output = vector("list", length = 4)

for(i in 1:4){
  
  output[[i]] = mean_and_sd(list_normals[[i]])
  
}

output
```

Use `map` to do this thing.

```{r}
map(list_normals, mean_and_sd)

map(list_normals, median)
```

check out some `map` variants

```{r}
map_dfr(list_normals, mean_and_sd, .id = "sample") # results in dataframe

map_dbl(list_normals, median) # results in a list of numbers
```

## List columns

Try to put my list into a dataframe!
```{r}
listcol_df = 
  tibble(
    name = c("a", "b", "c", "d"),
    sample = list_normals
  )

listcol_df
```

Did this realy work??

```{r}
pull(listcol_df, name)
pull(listcol_df, sample)
```

Can I apply `mean_and_sd`?
```{r}
mean_and_sd(pull(listcol_df, sample)[[1]])
mean_and_sd(pull(listcol_df, sample)[[2]])
mean_and_sd(pull(listcol_df, sample)[[3]])
mean_and_sd(pull(listcol_df, sample)[[4]])
```

iterate using `map`

```{r}
map(pull(listcol_df, sample), mean_and_sd)
```

adding a column

```{r}
listcol_df =
  listcol_df |> 
  mutate(
    summary = map(sample, mean_and_sd)
  )

pull(listcol_df, summary)
```

```{r}
listcol_df |> 
  select(-sample) |> 
  unnest(summary)
```



```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

nsduh_html = read_html(nsduh_url)


nsduh_import = function(html, table_num){
  
  data = 
    html |> 
    html_table() |> 
    nth(table_num) |>
    slice(-1) |> 
    select(-contains("P Value")) |>
    pivot_longer(
    -State,
    names_to = "age_year", 
    values_to = "percent") |>
    separate(age_year, into = c("age", "year"), sep = "\\(") |>
    mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent)) |>
    filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
    
  return(data)
  
}

nsduh_import(nsduh_html, table_num = 1)
nsduh_import(nsduh_html, table_num = 2)
nsduh_import(nsduh_html, table_num = 3)
```

Try this with a `for` loop

```{r}
output = vector("list", length =3)

for (i in 1:3){
  output[[i]] =   nsduh_import(html = nsduh_html, i)
}
```

Do this with `map`

```{r}
map(1:3, nsduh_import, html = nsduh_html)
```

Do this al in a dataframe

```{r}
nsduh_df = 
  tibble(
    name = c("major year", "major month", "major first"),
    number = 1:3
  ) |> 
  mutate(
    table = map(number, nsduh_import, html = nsduh_html)
  ) |> 
  unnest(table)
```

## Look at weather data

```{r}
library(p8105.datasets)
data("weather_df")
```

```{r}
weather_df |> 
  filter(name == "CentralPark_NY") |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point() +
  geom_smooth(method = "lm")
```

Let's do a regression

```{r}
weather_df |> 
  filter(name == "CentralPark_NY") |> 
  lm(tmax ~tmin, data = _)

weather_df |> 
  filter(name == "Molokai_HI") |> 
  lm(tmax ~tmin, data = _)

weather_df |> 
  filter(name == "Waterhole_WA") |> 
  lm(tmax ~tmin, data = _)
```

Let's iterate differently

```{r}
weather_nest =
  weather_df |> 
  nest(data = date:tmin)
```

```{r}
lm(tmax ~ tmin, data = pull(weather_nest, data)[[1]])
lm(tmax ~ tmin, data = pull(weather_nest, data)[[2]])
lm(tmax ~ tmin, data = pull(weather_nest, data)[[3]])
```

Do this using `map`

```{r}
weather_lm = function(df){
  lm(tmax ~ tmin, data = df)
}

```

```{r}
map(pull(weather_nest, data),weather_lm)
```

```{r}
weather_nest |> 
  mutate(
    lm_fits = map(data, weather_lm)
  )
```

